---
title: "FIFA Rating Predictions"
output:
  pdf_document: default
  html_notebook: default
---

# Fetching from DB

Start by connection to the database

```{r}
library(DBI)
con <- dbConnect(RPostgres::Postgres(),
                 dbname = 'postgres', 
                 host = 'localhost',
                 port = 5432,
                 user = 'postgres',
                 password = 'password')
```

Matches players stats to ratings from the previous year and their current year final rating

```{r}
res <- dbSendQuery(con, ("SELECT p.*, ps.*, curr.rating as currrating, prevr.*
FROM PLAYER p
JOIN PLAYER_STATS ps ON p.player_id = ps.player_id
JOIN PLAYER_RATINGS prevr ON p.player_id = prevr.player_id
JOIN PLAYER_RATINGS curr ON p.player_id = curr.player_id
WHERE ps.season = prevr.season AND ps.season = curr.season-1;"))



tab = dbFetch(res)
dbClearResult(res)
```

```{r}
tab
```

# Modeling

Data is separated into two sets for regression and classification and split into train and test sets

```{r}
library(tidyverse)
library(glmnet)
tab[is.na(tab)] <- 0
data_reg <- tab
data_class <- tab
data_class$currrating <- as.factor(data_class$currrating)


set.seed(100)
train_reg <- data_reg %>% sample_frac(size = 0.7, fac=currrating)
test_reg <- data_reg %>% setdiff(train_reg)
train_class <- data_class %>% sample_frac(size = 0.7, fac=currrating)
test_class <- data_class %>% setdiff(train_class)
```

While different formulas were tested, I settled on using all variables for the final prediction

```{r}
library(rpart)
library(rpart.plot)
#form <- as.formula(currrating~apps+minutes+goals+xG+assists+xA+shots+key_passes+npg+npxG+xGChain+xGBuildup+rating+pac+dri+pas+def+phy)
form <- as.formula(currrating~.)
```

Random forest is used for both regression and classification

```{r}
library(randomForest)
reg_rf <- randomForest(form,train_reg,ntree=1000,importance=TRUE)
class_rf <- randomForest(form,train_class,ntree=1000,importance=TRUE)
```

# Results

```{r}
library(caret)
p1_reg <- predict(reg_rf, train_reg)
print(mean((p1_reg-train_reg$currrating)^2))
```

```{r}
p1_class <- predict(class_rf, train_class)
confusionMatrix(p1_class, train_class$currrating)
```

```{r}
p2_reg <- predict(reg_rf, test_reg)
print(mean((p2_reg-test_reg$currrating)^2))
```

```{r}
p2_class <- predict(class_rf, test_class)
confusionMatrix(p2_class, test_class$currrating)
```

# Getting Predictions

```{r}
res2 <- dbSendQuery(con, ("SELECT p.*, ps.*, prevr.*
FROM PLAYER p
JOIN PLAYER_STATS ps ON p.player_id = ps.player_id
JOIN PLAYER_RATINGS prevr ON p.player_id = prevr.player_id
WHERE ps.season = prevr.season and ps.season = 23;"))



tab2 = dbFetch(res2)
dbClearResult(res2)
```

```{r}
tab2
```

```{r}
data2 <- tab2
data2 <- data2 %>% rename('player_id..27' = 'player_id..26')
data2 <- data2 %>% rename('season..28' = 'season..27')
data2[is.na(data2)] <- 0
```

```{r}
library(modelr)
final <- data2 %>% spread_predictions(reg_rf, class_rf) %>% 
  select(player_name_futbin, rating, reg_rf, class_rf) %>%
  arrange(-rating) %>%
  rename('name' = 'player_name_futbin', 
         'FIFA 23' = 'rating', 'Regression' = 'reg_rf', 
         'Classification' = 'class_rf')
final
```

```{r}
write.csv(final, "predictions.csv", row.names=FALSE)
```
